IFC Inside: A General Approach to Retrofitting Languages with Dynamic Information Flow Control

### 2014-03-02 paper/intro.tex
John's stuff
Signed-off-by: Edward Z. Yang <ezyang@cs.stanford.edu>

A defining aspect of web platforms is the combinability of
components of differing providence.  This pattern shows up in browser
extensions (which run JavaScript code alongside third-party website
code), web applications (which include possibly untrusted third-party
libraries), and mash-up websites (which combine code and data from
websites which may not even be aware of each other's existence.)  While
this combinability has offered great flexibility, it also poses a set of
unique security challenges: in particular, maintaining the privacy of
data in these settings is difficult, if not impossible.

Information flow control (IFC) is a promising technique for addressing these
problems by tracking the flow of sensitive data through a system. Untrusted
code is confined so that it cannot exfiltrate data, except as
per an information flow policy.  Significant research has been devoted to
adding various forms of IFC to different kinds of programming languages
and systems.  In the context of the web, however, there is a strong
motivation to preserve the language JavaScript, and retrofit dynamic information
flow control on top of it.  A key challenge in the design of these
systems is ensuring they can be implemented efficiently: when the
semantics specify that information flow must be tracked at a
fine-grained level, e.g., per value in the system,  it can lead to
impractical runtime overheads even for ordinary operations.

One alternative is a \textit{coarse-grained} approach to IFC.  Here,
computational units are associated with a floating label which protects
all data in this unit.  A recent system named SWAPI adds IFC to the
browser by identifying \emph{web workers} as an existing computational
unit.  Communication between workers can efficiently be mediated in an
IFC aware way to ensure non-interference, where sensitive workers cannot
influence the computations in less sensitive workers.
This coarse-grained approach provides a number
of advantages: (1) reduced runtime overhead because checks need only
be performed at isolation boundaries, (2) minimal changes to an
existing programming language when adding IFC, which (3) allows
the reuse of existing programs.  Finally, (4) the simplicity
of a single security label per computational unit facilitates
understanding and easy reasoning about the correctness of the
system, in particular without having to reason about all the
technical details of the semantics of the programming language
at hand.

In fact, the simplicity of this coarse-grained approach
to IFC allows us to generalize it to a large class of programming
languages.  In this paper, we identify the key concepts of coarse-grained
dynamic IFC and show how these can be combined with \emph{any} programming
language.  We achieve this
by using an embedding that combines the operational semantics
of the core IFC calculus and the target language.
At the core of our IFC system we use a sandboxing operation that allows
the execution of potentially sensitive computations in isolation,
while communicating through clearly specified interfaces with other
sandboxed operations.
We formally state and prove the non-interference guarantees that our
system gives, independently of the choice of the target language.

One challenge in the design of our system is that the semantics for the
retrofitted IFC language may not reflect the actual implementation in a
particular choice of target language.  We partially address this issue
by characterizing isomorphisms between the operational semantics of our
defined language and a concrete implementation, showing that if this
relationship holds, then noninterference in the abstract specification
carries over to the concrete specification.

Our key contributions can be summarized as follows:
\begin{itemize}
  \item We give formal semantics for a core IFC language that
      captures the essence of coarse-grained information flow control.
  We then show how a large class of target languages can be combined
  with this IFC language and prove that the resulting system provides
  non-interference. (Section~\ref{sec:retrofit})
%  \item We show how to extend the Matthews-Findler method from just
%  languages that are not simply expressions evaluating to values,
%  but may be collections of threads executing nondeterministically.
  \item We provide a proof technique to show the non-interference
  of a concrete semantics for a potentially optimized IFC language
  by means of an isomorphism, and show a class of restrictions on
  the IFC calculus that preserves non-interference. (Section~\ref{sec:formal})
  \item We sketch how the core IFC language can be enriched with
  more advanced concepts such as labeled values, privileges or a
  notation of clearance. (Section~\ref{sec:extensions})
  \item We connect our formal semantics to real implementations of
  coarse-grained IFC systems for Javascript and Haskell, and explain
  how our system could be used to enforce IFC in C. (Section~\ref{sec:real})
  \item We informally connect our results to a monadic interpretation
      of IFC which guided our design. (Section~\ref{sec:monad})
\end{itemize}

>>>>

Modern web content is rendered using a potentially large number of
different components with differing provenance.
Disparate and untrusting components may arise from browser
extensions (whose JavaScript code runs alongside website
code), web applications (with possibly untrusted third-party
libraries), and mashups (which combine code and data from
websites that may not even be aware of each other's existence.)
While just-in-time combination of untrusting components
offers great flexibility, it also poses complex security challenges.
In particular, maintaining data privacy in the face of malicious
extensions, libraries, and mashup components has been difficult,
and in general effectively impossible.

Information flow control (IFC) is a promising technique
that provides security
by tracking the flow of sensitive data through a system.
Untrusted code is confined so that it cannot exfiltrate data, except as
per an information flow policy.  Significant research has been devoted to
adding various forms of IFC to different kinds of programming languages
and systems.  In the context of the web, however, there is a strong
motivation to preserve the language JavaScript, and retrofit dynamic information
flow control on top of it.  A key challenge in dynamic information
flow control is implementation efficiency: when
information flow is tracked at a fine-grained level,
e.g., per value in the system, systems may incur
impractical runtime overhead.

In the alternative \textit{coarse-grained} approach to IFC,
a system is divided into relatively course computational units,
each with a single label dictating its security policy.
Only
communication between isolated computational units must be tracked.
A recent system named SWAPI adds IFC to the
browser by identifying \emph{web workers} as an existing computational
unit.  Communication between workers is mediated using
IFC to ensure non-interference, preventing sensitive workers from
influencing the computation of less sensitive workers.
This coarse-grained approach provides a number of advantages:
(1) reduced runtime overhead because checks need only
be performed at isolation boundaries, (2) minimal changes to an
existing programming language when adding IFC, which (3) allows
the reuse of existing programs.  Finally, (4) associating
a single security label with the entire computational unit simplifies
understanding and easy reasoning about the correctness of the
system, without information-flow reasoning about the
technical details of the semantics of the programming language.

A further benefit of the coarse-grained approach
is that a single IFC design and semantics can be generalized
to a large class of programming languages.
Leveraging this insight, we define a general form of coarse-grained
dynamic IFC and show how IFC can be combined with \emph{any} programming
language conforming to a general semantic framework.
At the core of our IFC system is a sandboxing construct that isolates
potentially sensitive computations, allowing them to
communicate only through clearly specified interfaces.
Based on a framework for combining the operational semantics
of our core IFC calculus with the operational semantics of any target language,
we formally state and prove non-interference guarantees that are
independent of the choice of specific target language.

One challenge in the design of our system is that the semantics for the
retrofitted IFC language may not accurately correspond to the optimized
implementation details of a particular target language.
We discuss and partially address this issue
by characterizing isomorphisms between the operational semantics of our
defined language and a concrete implementation, showing that if this
relationship holds, then noninterference in the abstract specification
carries over to the concrete specification.

Our contributions can be summarized as follows:
\begin{itemize}
  \item We give formal semantics for a core coarse-grained
  information flow control language.
  We then show how a large class of target languages can be combined
  with this IFC language and prove that the result provides
  non-interference. (Section~\ref{sec:retrofit})
%  \item We show how to extend the Matthews-Findler method from just
%  languages that are not simply expressions evaluating to values,
%  but may be collections of threads executing nondeterministically.
  \item We provide a proof technique to show the non-interference
  of a concrete semantics for a potentially optimized IFC language
  by means of an isomorphism, and show a class of restrictions on
  the IFC language that preserves non-interference. (Section~\ref{sec:formal})
  \item We briefly describe ways to enrich the core IFC language with
  more advanced concepts such as labeled values, privileges, or a
  notation of clearance. (Section~\ref{sec:extensions})
  \item We connect our formal semantics to real implementations of
  coarse-grained IFC systems for Javascript and Haskell, and explain
  how our system could be used to enforce IFC in C. (Section~\ref{sec:real})
  \item We informally relate our results to a monadic interpretation
      of IFC which guided our design. (Section~\ref{sec:monad})     
\end{itemize}
